### 2014-08-14 Studies/2011_San_Francisco/Report/San_Francisco_Fire.tex
San Francisco: Edits to report (fuel section, timeline); comment out pending sections

The analysis of the simulation results is focused on the interior conditions on the second floor before and after the interior door failure. Based on the goals of the analysis and the information collected, several engineering approximations were made. For example, the fire exhibited signs of under-ventilated conditions when firefighters arrived~\cite{NIOSH:Bowyer2}; however, the exact ignition time, fire growth rate, and heat release rate (HRR) of the fire were not known. The ignition time and growth rate were estimated to reduce the amount of required calculation time while adhering to documented observations. The HRR of the fire was estimated using documentation of the potential fuel load within the structure.

In reality, the quantities associated with these input parameters are not fixed values; rather, a model input parameter can be thought of as a point estimate from a distribution of possible input parameters with some associated amount of uncertainty. Any change in an input parameter (such as the HRR) for a given scenario would result in a change in the output quantity (such as the hot gas layer (HGL) temperature). For example, according to the McCaffrey, Quintiere, and Harkleroad~\cite{SFPE:Walton} empirical correlation, the HGL temperature in a well-ventilated compartment fire is proportional to the HRR raised to the two-thirds power. Following this relationship, a 7.5~\% increase in the HRR would result in a 5~\% increase in the HGL temperature~\cite{NUREG_1824_Sup_1}. More detailed discussion on the propagation of parameter uncertainty in fire models is available in a validation study that was sponsored by the U.S. Nuclear Regulatory Commission~\cite{NUREG_1824_Sup_1}. The following subsections describe the inputs that were used to develop the simulation, including any approximations or assumptions that were made.

@@

The structure involved in this fire incident was a 2~\sfrac{1}{2}~story residential building with balloon frame construction and a gable roof.\footnote{Balloon frame construction is a method of framing where a continuous stud runs from the bottom horizontal member to the top plate. Intermediate floors are built directly off of the studs. One potential issue with balloon frame construction is that without fire stops there is void space for fire to potentially travel between the floors within the wall cavity.} The structure had a 17.15~m (56~ft) by 6.15~m (20~ft) footprint, as shown in Fig.~\ref{fig:simp_geom}. This results in a floor area of approximately 105~m$^2$ (1100~ft$^2$) per floor and a main living area of approximately 110~m$^2$ (2200~ft$^2$). The interior walls were constructed from nominal lumber covered by 12.7~mm (0.5~in) gypsum board~\cite{NIOSH:Bowyer2}. Additionally, there was a half-story attic that contained the structure's hot water heater. Three sets of stairs were present in the structure: one front stairwell that connected the first and second levels (Stair 1 in Fig.~\ref{fig:simp_geom}), one rear stairwell that connected the first and second levels (Stair 2 in Fig.~\ref{fig:simp_geom}) and a second rear set of stairs which provided access to the attic (Stair 3 in Fig.~\ref{fig:simp_geom}). Entry to the first floor of the structure can be made either in the front or rear, as shown in Fig.~\ref{fig:simp_geom}.

The fire, which is described in Section~\ref{sec:fire}, started in the attic and spread to the rear porch. Discussion of the fire's impact on the interior conditions is provided in Section~\ref{sec:model_results}. Figure~\ref{fig:alpha_ex} shows the exterior of the front and rear of the structure after the incident. Note that the majority of the damage occurred in the attic and rear porch on the second floor. Fully dimensioned drawings of the interior of the first and second floors are shown in Figs.~\ref{fig:first_floor} and \ref{fig:second_floor}, respectively.

@@

To estimate the type of fuel, fire size, and structure ventilation, on-scene videos and post-incident reports were used. Based on these sources of information, the FDS model includes two source fire locations: an initial fire in the attic and a secondary fire on the second floor porch landing. The attic and rear porch primarily consisted of wood fuel: the rear porch was enclosed with wood planks covered with foam board and vinyl siding, the roof consisted of asphalt shingles over wood planking, and the attic and porch floors were composed of oriented strand board.

Wood fuel can be represented by the chemical formula, $\C\H_{1.7}\O_{0.74}\N_{0.002}$, with specified yields of carbon monoxide ($y_{\mathrm{CO}}=0.004 \; \mathrm{kg}/\mathrm{kg}$) and soot ($y_{\mathrm{C}}=0.015 \; \mathrm{kg}/\mathrm{kg}$)~\cite{SFPE:Tewarson}. The product yields are expressed in terms of the amount of carbon monoxide or soot emitted per unit mass of fuel consumed (kg/kg) and can be found in the Society of Fire Protection Engineers (SFPE) Handbook~\cite{SFPE:Tewarson}. A balanced chemical reaction for wood combustion can written as:

\begin{multline}
\C\H_{1.7}\O_{0.74}\N_{0.002} + 4.89(0.208\,\O_{2} + 0.783\,\N_{2} + 0.387\text{\sc{e}-}3\,\C\O_{2} + 0.834\text{\sc{e}-}2\,\H_{2}\O) \\ 
\rightarrow 5.84(0.719\text{\sc{e}-}3\,\C\O + 0.168\,\C\O_{2} + 0.155\,\H_{2}\O + 0.669\,\N_{2} + 0.693\text{\sc{e}-}2\,\C)
\label{eq:wood_comb}
\end{multline}

The value for the heat of combustion of wood used in this simulation was~16,400 kJ/kg, based on data provided in the SFPE Handbook~\cite{SFPE:Tewarson}. The heat of combustion quantifies the amount of energy per unit mass of the fuel. The following text defines Eq.~\ref{eq:wood_comb} for an FDS input file with the desired fuel and reaction properties discussed above:

\begin{lstlisting}
&SPEC ID='WOOD', FORMULA='CH1.7O0.74N0.002' /

&REAC ID = 'wood' 
    FUEL = 'WOOD', 
    HEAT_OF_COMBUSTION = 16400,
    SOOT_YIELD = 0.015,
    CO_YIELD = 0.004/
\end{lstlisting}

@@

To estimate the heat release rate per unit area (HRRPUA) of wood, Babrauskas and Grayson~\cite{babrauskas1990} conducted experiments in a cone calorimeter\footnote{The cone calorimeter is an experimental apparatus used to gather data about ignition time, mass loss, combustion products, and heat release rate among other properties associated with burning small samples of materials~\cite{ASTM:E1355}.} to determine the 5-min average of the HRRPUA for several different types of wood over a range of radiant heat fluxes. The results of that study indicate that the HRRPUA for different types of wood is approximately 50~kW/m$^2$.

To calculate the total fire size, the appropriate burning area (exposed wood area) of the attic and second floor porch enclosure needs to be determined. Based on the floor plan of the structure, the underside of the roof was approximately 134~m$^2$ (1453~ft$^2$) with 7~m$^2$ (75~ft$^2$) walls on the front and rear sides. Additionally, there was 105~m$^2$ (1130~ft$^2$) of floor area in the attic. Examination of post-incident images indicated that the floor immediately over the rear porch burned away. That area was considered to be part of the porch fire. Based on post-incident damage, approximately 50~\% of the remaining floor area and 90~\% of the roof was considered to be fuel for the attic fire. To estimate the total attic fire size, the underside roof area (120~m$^2$), two end walls (14~m$^2$), and the floor area (43~m$^2$ or 463~ft$^2$) were included. Applying an HRRPUA of wood of 50~kW/m$^2$ results in a potential peak fire size of 9~MW. To estimate the size of the porch fire, the four walls of the porch enclosure (42~m$^2$ or 452~ft$^2$) and the floor and ceiling of the porch (24~m$^2$ or 258 ft$^2$) were included, which results in a total porch area of 66~m$^2$ (689~ft$^2$). Using the same HRRPUA of wood of 50~kW/m$^2$, the potential peak fire size on the porch was approximately 3.3~MW. Therefore, the maximum fire size for this structure was estimated to be 12.3~MW. Since the time at which the fire started is not known relative to the notification of a fire or how fast the fire spread, several simplifications were made. The first assumption is that the attic fire was specified to start at the beginning of the simulation and increase to a value of 9~MW in 10~s. The porch fire was specified to start 40~s into the simulation and increase to a value of 3.3~MW in 10~s. The 10-s ramp up time was used to prevent abrupt changes in the velocity at the fire boundary condition in FDS. The second assumption is that the sources of the gas phase wood fuel have constant areas at fixed locations (2 sources for the attic, 1 source for the porch). The following lines define the source fires in the FDS input file:

\begin{lstlisting}
&SURF ID='FIRE', HRRPUA=2250., COLOR='ORANGE', TAU_Q=10 /
&VENT XB=-13.5,-13,1,5,6.16,6.16, SURF_ID='FIRE' / 
&VENT XB=-14.5,-14,1,5,6.16,6.16, SURF_ID='FIRE' / 
&VENT XB=-16.5,-15.5,1,2.5,3.14,3.14, SURF_ID='FIRE', DEVC_ID='SECOND FIRE'/ 
&DEVC XYZ = -17.2,2,0, ID='SECOND FIRE', QUANTITY='TIME', SETPOINT=40, INITIAL_STATE=.FALSE. /
\end{lstlisting}
Note that the coordinates here are unique for the vents are specific to the input files used for these simulations. The key point is that they are 2-D planes located at a specific area.

@@

While the fires were represented as fuel sources with a constant area (Section~\ref{sec:fire}), from a heat transfer perspective, it is important to define the material properties (density, thermal conductivity, and specific heat) to account for heat transfer and energy storage in the ceiling, walls, and floor. In this study, the material properties of gypsum board~\cite{WAKILI2007} were specified on the finished walls and ceilings on the main two floors of the structure, and the material properties of wood~\cite{Incropera:1} were specified on the unfinished portions of the attic and rear porch. The following lines define these materials in the FDS input file:

\begin{lstlisting}
&MATL ID = 'GYPSUM BOARD'
    FYI = 'Wakili - Journal of Fire Sciences 2007' 
    CONDUCTIVITY = 0.28
    SPECIFIC_HEAT = 1.0
    DENSITY = 810. /

&MATL ID = 'WOOD MAT'
    FYI = 'Incropera and DeWitt, Fundamentals of Heat and Mass Transfer'
    CONDUCTIVITY = 0.12
    SPECIFIC_HEAT = 1.38
    DENSITY = 510. / 
\end{lstlisting}

@@

The simulations in this study account for changes in the ventilation due to a combination of fire department operations (opening doors) and fire acting on the structure (fire breaching walls, breaking windows, etc). Two different ventilation scenarios were considered in this study. The first scenario involves a baseline simulation in which the ventilation times and ventilation areas represent our best understanding of the incident. The second scenario, or alternative simulation, does not include the firefighter opening the rear door to make entry to the structure and the subsequent exterior left wall fire penetration. The times of the ventilation changes are provided in Table~\ref{tab:ventilation_timeline}. Structure leakage has been shown to be important when modeling enclosure fires~\cite{beal2009}. The open front door at the start of the simulation and the vents (opening size and open time) created during the simulation are significantly larger than the total effective area of leakage of a structure of this type. Therefore, any leakage in the structure would have a negligible impact on the fluid mechanics within the structure and is not included in this study.

An unknown parameter in Table~\ref{tab:ventilation_timeline} that needed to be estimated was the time when the second floor interior/porch door failed (160~s into the simulation). The area of failure was also estimated based on the following observations. The bottom hinge from the door appeared to be closed throughout the fire due to the lack of fire damage and the lack of soot deposited on the interior faces of the hinge and remaining wood, as shown in Fig.~\ref{fig:door_frame}. It appears that the door collapsed and failed partially open. In the simulations, only the upper portion of the door was assumed to fail. While this time and area may not be exact, the takeaway from the analysis in the following sections should be impact of the door's failure on the interior conditions of the structure.

@@

For the simulation, a measure of how well the flow field is resolved can be estimated by using the non-dimensional expression $D^*/\dx$. Here, $D^*$ is the characteristic fire diameter, $\dx$ is the nominal size of a mesh cell, and $\dQ$ is the total heat release rate of the fire:
\begin{equation}
D^* = \left(
     \frac{\dQ}{\rho_\infty \, c_p \, T_\infty \, \sqrt{g}}
     \right)^\frac{2}{5} 
\label{eq:mesh}
\end{equation}
From the FDS User's Guide~\cite{FDS_Users_Guide}, the characteristic fire diameter is related to the characteristic fire size via the
relation $Q^* = (D^*/D)^{5/2}$. Here, $D$ is the physical diameter of the base of the fire specified in the simulation. Following Eq.~\ref{eq:mesh} and using a grid cell size of 10~cm, the characteristic fire diameter to cell size ($D^*/\dx$) ratio is 15.5 for the 3.3~MW porch fire and 23 for the 9~MW attic fire. Based on validation work performed for the U.S. Nuclear Regulatory Commission, $D^*/\dx$ values ranged between 4 and 16~\cite{NUREG_1824} and produced results that were adequate for engineering calculations. The grid resolution used in this model is within or exceeds typical engineering $D^*/\dx$ values.

As discussed in Section~\ref{sec:geometry}, the structure has dimensions of 17.15~m (56~ft) by 6.15~m (20~ft) and a total height of 8.56~m (28~ft). To ensure adequately resolved fluid flow in and out of the structure, the computational domain was extended beyond the volume of the structure. The total volume of the computational domain was 24~m (79~ft) by 10~m (33~ft) by 10~m (33~ft). A grid resolution of 10~cm (3.9~in) was used, which required 2.4 million computational cells. As a result, all of the input model geometry lengths, ventilation openings (holes, doors, and windows), and fire areas will snap to the nearest 10~cm. Figure~\ref{fig:geom_grid} shows the front of the structure rendered in Smokeview with the 10~cm mesh. The domain was divided into 16 equally sized meshes (each containing 150,000 grid cells) that could be processed in parallel, which reduced the amount of required calculation time to approximately 1.5~days. Figure~\ref{fig:mult_mesh} shows the entire structure within the computational domain that has been divided into multiple meshes. Table~\ref{tab:model_parameters} shows a summary of the model input parameters for the simulations that were conducted as part of this study.

@@

17:15:00       &              &  Fire is reported with heavy smoke coming from the               \\ 
               &              &  front and rear of the structure's attic.                        \\

@@

17:16:00       &              &  Dispatch for a ``Still'' alarm - two engine companies,          \\
               &              &  two truck companies, and a battalion chief.                     \\

@@

               &  0           &  FDS simulation starts. Attic fire ignites. Initial              \\
               &              &  ventilation through rear of structure. Front door open.         \\

@@

17:20:00       &  40          &  Attic windows break - heavy smoke out of front attic            \\
               &              &  window and heavy smoke and fire out of rear.                    \\
               &              &  Rear porch ignites in model.                                    \\

@@

               &  45          & Ventilation through wall penetration on rear of structure.       \\

@@

               &  55          & Ventilation through wall penetration on left side of structure.  \\

@@

17:21:00       &  100         & E123 captain and firefighter on second floor                     \\
               &              & with 1\sfrac{3}{4}-inch hoseline.                                \\

@@

17:21:30       &  130         & First floor exterior porch door opened.                          \\
               &              & Porch fire ignites in FDS model.                                 \\

@@

               &  135         & Ventilation through left side of structure.                      \\

@@

               &  160         & Top of interior door fails in FDS model.                         \\

@@

17:23:00       &  220         & Heavy fire report in rear stairwell and enclosed porch.          \\
               &              & No response from E123 captain or firefighter.                    \\

@@

17:24:00       &              & Water applied to attic from the rear exterior position.          \\

@@

17:25:00       &              & SQ5 makes entry through front of structure.                      \\
17:27:00       &  300         & After hearing a Mayday call from SQ5, IC calls a Mayday.         \\
               &              & Simulation ends.                                                 \\
17:29:00       &              & E123 captain carried out of structure.                           \\

@@

The heat release rate output data from FDS quantifies the simulated rate of energy release in the simulation. Figure~\ref{fig:hrr} shows the prescribed HRR versus time values based on the prescribed fires along with the FDS results for the two ventilation scenarios that were described in Section~\ref{sec:ventilation}. The two vertical lines represent the time at which a firefighter opened the back door of the structure ($t$ = 130 s) and the time at which the second floor interior door was estimated to fail ($t$ = 160 s). Due to a lack of information about the time of fire ignition or how fast the fire grew, these parameters needed to be estimated.  The growth rate shown in Fig.~\ref{fig:hrr} (10~MW in approximately 20~s) is fast because the incipient stage of the fire growth was not considered in this study. The maximum size of the fire used in this study was selected such that the conditions would become ventilation limited quickly to minimize computational time while aligning with the timeline observations of video and picture evidence from the fire. Using this approach, the accuracy of heat transfer to the walls of the structure may be sacrificed; however, all of the window, door, and wall failures were driven by the event timeline rather than failure temperature criteria.

@@

In Fig.~\ref{fig:hrr}, the solid line represents the baseline FDS simulation (this follows the noted door openings from the incident). The dashed line represents an alternative series of events in which the rear door does not get opened and the subsequent breach of the rear porch wall on the left side of the structure does not occur. The dash-dotted line represents the prescribed (or nominal) HRR that was input into FDS. As shown in Fig.~\ref{fig:hrr}, the HRR output from FDS is less than the prescribed HRR because the fire in the structure becomes ventilation limited. The maximum HRR value is slightly less than 9~MW before the fire becomes ventilation limited and begins to decay. The HRR is reduced to 5~MW as the oxygen inside the structure is consumed. There is an insufficient amount of oxygen in the structure to burn all of the fuel, therefore combustion occurs locally at the vents where fresh air is available. Figure~\ref{fig:smv_ext_fire} shows flaming combustion through the failed window on the rear side of the structure and attic breach just after the second fire ignites at 40~s. After additional wall breaches occur at 45~s and 55~s (see Table~\ref{tab:ventilation_timeline}), the abrupt change in ventilation causes the HRR to increase to 9~MW before it decreases to approximately 6~MW. The HRR then increases towards a sustainable steady-state value.

Figure~\ref{fig:hrr} indicates that the HRRs for both of the ventilation scenarios in FDS are identical until 130~s (HRR $\approx$ 7~MW). At 130~s, the rear door is opened in the baseline simulation. At 135~s, there is an additional breach on the left side of the structure (Table~\ref{tab:ventilation_timeline}). The open door and breach act as additional pressure releases for the structure. Again, the abrupt change in ventilation causes the HRR to increase from approximately 7~MW to 9.7~MW at 139~s. At the time of the door failure (160~s), the HRR in both cases was approximately 9~MW because of the increased amount of ventilation. While there might be some locally different burning characteristics between the baseline simulation and the alternative simulation, the interior conditions in both cases presented a significant hazard to firefighters inside of the structure. Therefore, the analysis will focus on the results of the baseline simulation.

>>>>

% The analysis of the simulation results is focused on the interior conditions on the second floor before and after the interior door failure. Based on the goals of the analysis and the information collected, several engineering approximations were made. For example, the fire exhibited signs of under-ventilated conditions when firefighters arrived~\cite{NIOSH:Bowyer2}; however, the exact ignition time, fire growth rate, and heat release rate (HRR) of the fire were not known. The ignition time and growth rate were estimated to reduce the amount of required calculation time while adhering to documented observations. The HRR of the fire was estimated using documentation of the potential fuel load within the structure.

% In reality, the quantities associated with these input parameters are not fixed values; rather, a model input parameter can be thought of as a point estimate from a distribution of possible input parameters with some associated amount of uncertainty. Any change in an input parameter (such as the HRR) for a given scenario would result in a change in the output quantity (such as the hot gas layer (HGL) temperature). For example, according to the McCaffrey, Quintiere, and Harkleroad~\cite{SFPE:Walton} empirical correlation, the HGL temperature in a well-ventilated compartment fire is proportional to the HRR raised to the two-thirds power. Following this relationship, a 7.5~\% increase in the HRR would result in a 5~\% increase in the HGL temperature~\cite{NUREG_1824_Sup_1}. More detailed discussion on the propagation of parameter uncertainty in fire models is available in a validation study that was sponsored by the U.S. Nuclear Regulatory Commission~\cite{NUREG_1824_Sup_1}. The following subsections describe the inputs that were used to develop the simulation, including any approximations or assumptions that were made.

@@

% The structure involved in this fire incident was a 2~\sfrac{1}{2}~story residential building with balloon frame construction and a gable roof.\footnote{Balloon frame construction is a method of framing where a continuous stud runs from the bottom horizontal member to the top plate. Intermediate floors are built directly off of the studs. One potential issue with balloon frame construction is that without fire stops there is void space for fire to potentially travel between the floors within the wall cavity.} The structure had a 17.15~m (56~ft) by 6.15~m (20~ft) footprint, as shown in Fig.~\ref{fig:simp_geom}. This results in a floor area of approximately 105~m$^2$ (1100~ft$^2$) per floor and a main living area of approximately 110~m$^2$ (2200~ft$^2$). The interior walls were constructed from nominal lumber covered by 12.7~mm (0.5~in) gypsum board~\cite{NIOSH:Bowyer2}. Additionally, there was a half-story attic that contained the structure's hot water heater. Three sets of stairs were present in the structure: one front stairwell that connected the first and second levels (Stair 1 in Fig.~\ref{fig:simp_geom}), one rear stairwell that connected the first and second levels (Stair 2 in Fig.~\ref{fig:simp_geom}) and a second rear set of stairs which provided access to the attic (Stair 3 in Fig.~\ref{fig:simp_geom}). Entry to the first floor of the structure can be made either in the front or rear, as shown in Fig.~\ref{fig:simp_geom}.

% The fire, which is described in Section~\ref{sec:fire}, started in the attic and spread to the rear porch. Discussion of the fire's impact on the interior conditions is provided in Section~\ref{sec:model_results}. Figure~\ref{fig:alpha_ex} shows the exterior of the front and rear of the structure after the incident. Note that the majority of the damage occurred in the attic and rear porch on the second floor. Fully dimensioned drawings of the interior of the first and second floors are shown in Figs.~\ref{fig:first_floor} and \ref{fig:second_floor}, respectively.

@@

To estimate the type of fuel, fire size, and structure ventilation, on-scene videos and post-incident reports were used. Based on these sources of information, the FDS model includes two source fires: an initial couch fire in the basement near the rear side of the structure, and a secondary flashover fire that spanned the entire basement floor area. The basement consisted of a finished living area with a large utilities area and contained primarily upholstered polyurethane furniture fuel items.

Polyurethane fuel can be represented by the chemical formula, $\C\H_{1.8}\O_{0.30}\N_{0.05}$, with specified product yields of soot ($y_{\mathrm{s}}=0.131 \; \mathrm{kg}/\mathrm{kg}$) and carbon monoxide ($y_{\mathrm{CO}}=0.010 \; \mathrm{kg}/\mathrm{kg}$)~\cite{SFPE:Tewarson}. The product yields are expressed in terms of the amount soot or carbon monoxide emitted per unit mass of fuel consumed~(kg/kg) and can be found in the Society of Fire Protection Engineers (SFPE) Handbook~\cite{SFPE:Tewarson}. A balanced chemical reaction for polyurethane combustion can be written as:

\begin{multline}
\C\H_{1.8}\O_{0.30}\N_{0.05} + 5.22(0.208\,\O_{2} + 0.783\,\N_{2} + 8.34 \times 10^{-3}\,\H_{2}\O + 3.87 \times 10^{-4}\,\C\O_{2}) \\
\rightarrow 6.06(0.129\,\C\O_{2} + 0.156\,\H_{2}\O + 0.001\,\C\O + 0.035\,\C + 0.679\,\N_{2})
\label{eq:fuel_comb}
\end{multline}

The value for the heat of combustion of polyurethane used in this simulation was~26,200 kJ/kg, based on data provided in the SFPE Handbook~\cite{SFPE:Tewarson}. The heat of combustion quantifies the amount of energy per unit mass of the fuel. The following text defines Eq.~\ref{eq:fuel_comb} for an FDS input file with the desired fuel and reaction properties discussed above:

\begin{lstlisting}
&SPEC ID='POLYURETHANE', FORMULA='CH1.8O0.30N0.05' /

&REAC ID = 'POLYURETHANE' 
    FUEL = 'POLYURETHANE', 
    HEAT_OF_COMBUSTION = 26200,
    SOOT_YIELD = 0.131,
    CO_YIELD = 0.010 /
\end{lstlisting}

@@

% To estimate the heat release rate per unit area (HRRPUA) of wood, Babrauskas and Grayson~\cite{babrauskas1990} conducted experiments in a cone calorimeter\footnote{The cone calorimeter is an experimental apparatus used to gather data about ignition time, mass loss, combustion products, and heat release rate among other properties associated with burning small samples of materials~\cite{ASTM:E1355}.} to determine the 5-min average of the HRRPUA for several different types of wood over a range of radiant heat fluxes. The results of that study indicate that the HRRPUA for different types of wood is approximately 50~kW/m$^2$.

% To calculate the total fire size, the appropriate burning area (exposed wood area) of the attic and second floor porch enclosure needs to be determined. Based on the floor plan of the structure, the underside of the roof was approximately 134~m$^2$ (1453~ft$^2$) with 7~m$^2$ (75~ft$^2$) walls on the front and rear sides. Additionally, there was 105~m$^2$ (1130~ft$^2$) of floor area in the attic. Examination of post-incident images indicated that the floor immediately over the rear porch burned away. That area was considered to be part of the porch fire. Based on post-incident damage, approximately 50~\% of the remaining floor area and 90~\% of the roof was considered to be fuel for the attic fire. To estimate the total attic fire size, the underside roof area (120~m$^2$), two end walls (14~m$^2$), and the floor area (43~m$^2$ or 463~ft$^2$) were included. Applying an HRRPUA of wood of 50~kW/m$^2$ results in a potential peak fire size of 9~MW. To estimate the size of the porch fire, the four walls of the porch enclosure (42~m$^2$ or 452~ft$^2$) and the floor and ceiling of the porch (24~m$^2$ or 258 ft$^2$) were included, which results in a total porch area of 66~m$^2$ (689~ft$^2$). Using the same HRRPUA of wood of 50~kW/m$^2$, the potential peak fire size on the porch was approximately 3.3~MW. Therefore, the maximum fire size for this structure was estimated to be 12.3~MW. Since the time at which the fire started is not known relative to the notification of a fire or how fast the fire spread, several simplifications were made. The first assumption is that the attic fire was specified to start at the beginning of the simulation and increase to a value of 9~MW in 10~s. The porch fire was specified to start 40~s into the simulation and increase to a value of 3.3~MW in 10~s. The 10-s ramp up time was used to prevent abrupt changes in the velocity at the fire boundary condition in FDS. The second assumption is that the sources of the gas phase wood fuel have constant areas at fixed locations (2 sources for the attic, 1 source for the porch). The following lines define the source fires in the FDS input file:

\begin{lstlisting}
&SURF ID='BURNER', HRRPUA=0.25, COLOR='RED', RAMP_Q='Fire_Ramp' /
&VENT XB=4.2, 6.2, 15.2, 17.2, -2.8, -2.8, SURF_ID='BURNER' /

&SURF ID='FLASHOVER_BURNER', HRRPUA=270.563, COLOR='RED' /
&VENT XB=0.0, 8.4, 8.5, 15.1, -2.8, -2.8, SURF_ID='FLASHOVER_BURNER', DEVC_ID='TIMER_FLASHOVER' /
&DEVC ID='TIMER_FLASHOVER', QUANTITY='TIME', XYZ=0,0,0, INITIAL_STATE=.FALSE., SETPOINT=300 /
\end{lstlisting}
Note that the coordinates here are unique for the vents are specific to the input files used for these simulations. The key point is that they are 2-D planes located at a specific area.

@@

While the fires were represented as fuel sources with a constant area (Section~\ref{sec:fire}), from a heat transfer perspective, it is important to define the material properties (density, thermal conductivity, and specific heat) to account for heat transfer and energy storage in the ceiling, walls, and floor. In this study, the material properties of gypsum board~\cite{WAKILI2007} were specified on the finished walls and ceilings on the main two floors of the structure. The following lines define the gypsum material in the FDS input file:

\begin{lstlisting}
&MATL ID = 'GYPSUM BOARD'
    FYI = 'Wakili - Journal of Fire Sciences 2007' 
    CONDUCTIVITY = 0.28
    SPECIFIC_HEAT = 1.0
    DENSITY = 810. /
\end{lstlisting}

@@

% The simulations in this study account for changes in the ventilation due to a combination of fire department operations (opening doors) and fire acting on the structure (fire breaching walls, breaking windows, etc). Two different ventilation scenarios were considered in this study. The first scenario involves a baseline simulation in which the ventilation times and ventilation areas represent our best understanding of the incident. The second scenario, or alternative simulation, does not include the firefighter opening the rear door to make entry to the structure and the subsequent exterior left wall fire penetration. The times of the ventilation changes are provided in Table~\ref{tab:ventilation_timeline}. Structure leakage has been shown to be important when modeling enclosure fires~\cite{beal2009}. The open front door at the start of the simulation and the vents (opening size and open time) created during the simulation are significantly larger than the total effective area of leakage of a structure of this type. Therefore, any leakage in the structure would have a negligible impact on the fluid mechanics within the structure and is not included in this study.

% An unknown parameter in Table~\ref{tab:ventilation_timeline} that needed to be estimated was the time when the second floor interior/porch door failed (160~s into the simulation). The area of failure was also estimated based on the following observations. The bottom hinge from the door appeared to be closed throughout the fire due to the lack of fire damage and the lack of soot deposited on the interior faces of the hinge and remaining wood, as shown in Fig.~\ref{fig:door_frame}. It appears that the door collapsed and failed partially open. In the simulations, only the upper portion of the door was assumed to fail. While this time and area may not be exact, the takeaway from the analysis in the following sections should be impact of the door's failure on the interior conditions of the structure.

@@

For the simulation, a measure of how well the flow field is resolved can be estimated by using the non-dimensional expression $D^*/\dx$. Here, $D^*$ is the characteristic fire diameter, $\dx$ is the nominal size of a mesh cell, and $\dQ$ is the total heat release rate of the fire:
\begin{equation}
D^* = \left(
     \frac{\dQ}{\rho_\infty \, c_p \, T_\infty \, \sqrt{g}}
     \right)^\frac{2}{5} 
\label{eq:mesh}
\end{equation}
% From the FDS User's Guide~\cite{FDS_Users_Guide}, the characteristic fire diameter is related to the characteristic fire size via the relation $Q^* = (D^*/D)^{5/2}$. Here, $D$ is the physical diameter of the base of the fire specified in the simulation. Following Eq.~\ref{eq:mesh} and using a grid cell size of 10~cm, the characteristic fire diameter to cell size ($D^*/\dx$) ratio is 15.5 for the 3.3~MW porch fire and 23 for the 9~MW attic fire. Based on validation work performed for the U.S. Nuclear Regulatory Commission, $D^*/\dx$ values ranged between 4 and 16~\cite{NUREG_1824} and produced results that were adequate for engineering calculations. The grid resolution used in this model is within or exceeds typical engineering $D^*/\dx$ values.

% As discussed in Section~\ref{sec:geometry}, the structure has dimensions of 17.15~m (56~ft) by 6.15~m (20~ft) and a total height of 8.56~m (28~ft). To ensure adequately resolved fluid flow in and out of the structure, the computational domain was extended beyond the volume of the structure. The total volume of the computational domain was 24~m (79~ft) by 10~m (33~ft) by 10~m (33~ft). A grid resolution of 10~cm (3.9~in) was used, which required 2.4 million computational cells. As a result, all of the input model geometry lengths, ventilation openings (holes, doors, and windows), and fire areas will snap to the nearest 10~cm. Figure~\ref{fig:geom_grid} shows the front of the structure rendered in Smokeview with the 10~cm mesh. The domain was divided into 16 equally sized meshes (each containing 150,000 grid cells) that could be processed in parallel, which reduced the amount of required calculation time to approximately 1.5~days. Figure~\ref{fig:mult_mesh} shows the entire structure within the computational domain that has been divided into multiple meshes. Table~\ref{tab:model_parameters} shows a summary of the model input parameters for the simulations that were conducted as part of this study.

@@

10:45:00       &              &  Dispatch for a curtain fire due to an electrical short at a residential structure.  \\

@@

10:48:00       &              &  E26 arrives on scene, reports light smoke conditions, and makes entry into          \\
               &              &  the front door with a 1 3/4'' hoseline.                                             \\

@@

10:52:00       &  0           &  FDS simulation begins.                                                              \\

@@

10:54:00       &              &  E26 crew states that the fire is located below the first floor.                     \\

@@

10:56:00       &              &  BC9 observes smoke but no fire at the left rear corner of the structure.            \\

@@

10:58:00       &              &  Fire self-vents from the rear side of the structure when a glass door in the        \\
               &              &  basement fails.                                                                     \\

@@

10:59:00       &              &  BC9 forces open the basement door on the left side of the structure,                \\
               &              &  reports heavy fire and smoke, and requests a second hoseline.                       \\

@@

11:00:00       &              &  BC6 notices a severe change in conditions (heavy black smoke from garage).          \\

@@

11:01:00       &              &  Heavy fire and black smoke observed at rear of structure. Incident command          \\
               &              &  (IC) attempts to contact E26 several times via radio with no reply.                 \\

@@

11:02:00       &              &  E32 crew makes entry into the basement with a second hoseline.                      \\

@@

11:08:00       &              &  Two downed E26 firefighters are found on the first floor.                           \\

@@

11:09:00       &              &  The downed E26 firefighters are carried out of the structure.                       \\

@@

@@

% The heat release rate output data from FDS quantifies the simulated rate of energy release in the simulation. Figure~\ref{fig:hrr} shows the prescribed HRR versus time values based on the prescribed fires along with the FDS results for the two ventilation scenarios that were described in Section~\ref{sec:ventilation}. The two vertical lines represent the time at which a firefighter opened the back door of the structure ($t$ = 130 s) and the time at which the second floor interior door was estimated to fail ($t$ = 160 s). Due to a lack of information about the time of fire ignition or how fast the fire grew, these parameters needed to be estimated. The growth rate shown in Fig.~\ref{fig:hrr} (10~MW in approximately 20~s) is fast because the incipient stage of the fire growth was not considered in this study. The maximum size of the fire used in this study was selected such that the conditions would become ventilation limited quickly to minimize computational time while aligning with the timeline observations of video and picture evidence from the fire. Using this approach, the accuracy of heat transfer to the walls of the structure may be sacrificed; however, all of the window, door, and wall failures were driven by the event timeline rather than failure temperature criteria.

@@

% In Fig.~\ref{fig:hrr}, the solid line represents the baseline FDS simulation (this follows the noted door openings from the incident). The dashed line represents an alternative series of events in which the rear door does not get opened and the subsequent breach of the rear porch wall on the left side of the structure does not occur. The dash-dotted line represents the prescribed (or nominal) HRR that was input into FDS. As shown in Fig.~\ref{fig:hrr}, the HRR output from FDS is less than the prescribed HRR because the fire in the structure becomes ventilation limited. The maximum HRR value is slightly less than 9~MW before the fire becomes ventilation limited and begins to decay. The HRR is reduced to 5~MW as the oxygen inside the structure is consumed. There is an insufficient amount of oxygen in the structure to burn all of the fuel, therefore combustion occurs locally at the vents where fresh air is available. Figure~\ref{fig:smv_ext_fire} shows flaming combustion through the failed window on the rear side of the structure and attic breach just after the second fire ignites at 40~s. After additional wall breaches occur at 45~s and 55~s (see Table~\ref{tab:ventilation_timeline}), the abrupt change in ventilation causes the HRR to increase to 9~MW before it decreases to approximately 6~MW. The HRR then increases towards a sustainable steady-state value.

% Figure~\ref{fig:hrr} indicates that the HRRs for both of the ventilation scenarios in FDS are identical until 130~s (HRR $\approx$ 7~MW). At 130~s, the rear door is opened in the baseline simulation. At 135~s, there is an additional breach on the left side of the structure (Table~\ref{tab:ventilation_timeline}). The open door and breach act as additional pressure releases for the structure. Again, the abrupt change in ventilation causes the HRR to increase from approximately 7~MW to 9.7~MW at 139~s. At the time of the door failure (160~s), the HRR in both cases was approximately 9~MW because of the increased amount of ventilation. While there might be some locally different burning characteristics between the baseline simulation and the alternative simulation, the interior conditions in both cases presented a significant hazard to firefighters inside of the structure. Therefore, the analysis will focus on the results of the baseline simulation.
